PANCHANGE_data_preprocessing

### 2020-08-13 PANCHANGE_data_preprocessing2020-07-15.Rmd
Script running up until glad.gad

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
     # "Login ID", # ID
      startDate.ramp, #Date when participant signed up to ramp
      age_category = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
     # EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
     # demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
     # demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
    #  demographics.o_levelsgcses_or_equivalent, # level of education - gcse
     # demographics.cses_or_equivalent, # level of education - cse
     # demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# RAMP employment
## Read in data
```{r ramp emp cols}
if(RAMP == TRUE){
  ramp.emp.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_ramp.rds"))
  dim(ramp.emp.raw)
  colnames(ramp.emp.raw)
}
```

# Select variables
```{r ramp emp cleaning}
if(RAMP == TRUE){
  exclude_cols_emp <- c("ID",
                        "Sample")

ramp.emp.raw$externalDataReference <- ramp.emp.raw$`Login ID` #move Login ID to external data reference

ramp.emp.raw.id <- ramp.emp.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      highest_education_ramp = employment.what_is_your_highest_level_of_education,
      # EduYrs = employment.university_education_attend_school.txt, # Years in education - not in RAMP dataset
      employment.government_work_key_workers
    ) %>%
    add_numeric(., exclude = exclude_cols_emp) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(ramp.emp.raw.id)
  # Inspect colnames
  colnames(ramp.emp.raw.id)
  #Differences
  dim(ramp.emp.raw)[1]-dim(ramp.emp.raw.id)[1]
}
```

```{r ramp copy emp}
if(RAMP == TRUE){
  emp.raw.id <- ramp.emp.raw.id
}
```

@@

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

#recode the numeric version into a factor
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

glad.employment.raw.id %>%
  freq(highest_education)
}
```

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE) {
glad.employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

# Key worker status
```{r Key worker}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
#create numeric version of the variable
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    key_worker_numeric =
      case_when(
        employment.government_work_key_workers_numeric == "0" ~ 0,
        employment.government_work_key_workers_numeric > 0 ~ 1
      )
    )

#recode the numeric version into a factor
glad.employment.raw.id <- glad.employment.raw.id %>%
  mutate(
    key_worker =
      recode_factor(
        key_worker_numeric,
        `0` = "No key worker",
        `1` = "Key worker"
        )
    )

glad.employment.raw.id %>%
  freq(key_worker)
}
```

>>>>

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.ramp, #Date when participant signed up to ramp
      age_category = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
     # EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
     # demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
     # demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
    #  demographics.o_levelsgcses_or_equivalent, # level of education - gcse
     # demographics.cses_or_equivalent, # level of education - cse
     # demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# RAMP employment
## Read in data
```{r ramp emp cols}
if(RAMP == TRUE){
  ramp.employment.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_ramp.rds"))
  dim(ramp.employment.raw)
  colnames(ramp.employment.raw)
}
```

# Select variables
```{r ramp emp cleaning}
if(RAMP == TRUE){
  exclude_cols_employment <- c("ID",
                        "Sample",
                        "startDate.ramp")

ramp.employment.raw$externalDataReference <- ramp.employment.raw$`Login ID` #move Login ID to external data reference

ramp.employment.raw.id <- ramp.employment.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      Sample,
      ID,
      startDate.ramp,
      highest_education_ramp = employment.what_is_your_highest_level_of_education, #this is missing NVQ or equivalent which we have in GLAD-COPING
      # EduYrs = employment.university_education_attend_school.txt, # Years in education - not in RAMP dataset
      employment.government_work_key_workers
    ) %>%
    add_numeric(., exclude = exclude_cols_emp) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(ramp.employment.raw.id)
  # Inspect colnames
  colnames(ramp.employment.raw.id)
  #Differences
  dim(ramp.employment.raw)[1]-dim(ramp.employment.raw.id)[1]
}
```

```{r ramp copy emp}
if(RAMP == TRUE){
  employment.raw.id <- ramp.employment.raw.id
}
```

@@

```{r GLAD copy employment}
if(RAMP == GLAD){
  employment.raw.id <- glad.employment.raw.id
}
```

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

employment.raw.id %>%
  freq(highest_education)
}
```

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE) {
employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

# Key worker status
```{r Key worker}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
#create numeric version of the variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    key_worker_numeric =
      case_when(
        employment.government_work_key_workers_numeric == "0" ~ 0,
        employment.government_work_key_workers_numeric > 0 ~ 1
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    key_worker =
      recode_factor(
        key_worker_numeric,
        `0` = "No key worker",
        `1` = "Key worker"
        )
    )

employment.raw.id %>%
  freq(key_worker)
}
```
